---
title: "Palmer Penguins Analysis"
author: "Saran O."
format: html
editor: visual
execute: 
  keep-md: true
---

## Palmer Penguin Analysis

This is an analysis of Palmer's Penguin dataset.

## **Loading Packages and Datasets**

Here we will load the tidyverse package and penguins data.

```{r}
#Load the tidyverse
library(tidyverse)
library(kableExtra)

#Read the penguins_samp1 data file from github
penguins <- read_csv("https://raw.githubusercontent.com/mcduryea/Intro-to-Bioinformatics/main/data/penguins_samp1.csv")

#See the first six rows of the data we've read in to our notebook
penguins %>% 
  head(2) %>% 
  kable() %>% 
  kable_styling(c("striped", "hover"))
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## About Our Data

The data we are working with is a dataset on Penguins, which includes 8 features measured on 44 penguins. The features included are physiological features (like bill length, bill depth, flipper length, body mass, etc) ass well as other features like the year that the penguin was observed,the island the penguins was observed on, the sex of the penguin,and the species of the penguin.

## Interesting Questions to Ask

-   What is the average flipper length? What about for each species?

-   Are there more male or female penguins? What about per island or species?

-   What is the average body mass? What about by island? By species? By sex?

-   What is the ratio of bill length to bill depth for a penguin? What is the overall average of this metric? Does it change by species, sex, or island?

-   Does average body mass change year?

## Data Manipulation Tools and Strategies

We can look at individual columns in a data set or subsets of columns in a dataset. For example, if we are only interested in flipper length and species we can `select()` those columns

```{r}
penguins %>%
  select(species, body_mass_g)

```

If we want to `filter()` and only show certain rows, we can do that too.

```{r}
penguins %>%
  filter(species == "Chinstrap")

#We can also filter by nymerical variables
penguins %>%
  filter(body_mass_g >= 6000)

#We can also do both
 penguins %>%
   filter((body_mass_g >= 6000) | (island == "Torgersen"))
```

## Answering Our Questions

Most of our questions involve summarizing data, and perhaps summarizing over groups. We can summarize data using the `summarize()` functions, and data group using `group_by()`.

Let's find the average flipper length.

```{r}
#Overall average flipper length
penguins %>% 
  summarize(avg_flipper_length = mean(flipper_length_mm))

penguins %>%
  filter(species == "Gentoo") %>%
  summarize(avg_flipper_length = mean(flipper_length_mm))

#Grouped Average
penguins %>%
  group_by(species) %>%
  summarize(av_flipper_length = mean(flipper_length_mm))
```

How many of each species do we have?

```{r}
penguins %>%
  count(species)
```

How many of each sex are there? What about island or species?

```{r}
penguins %>%
 count(sex)

penguins %>%
  group_by(species) %>%
  count(sex)
```

We can `mutate()` to add new columns to our data set

```{r}
penguins_with_ratio <- penguins %>%
  mutate(bill_ltd_ratio = bill_length_mm / bill_depth_mm)

#Average Ratio
penguins %>%
  mutate(bill_ltd_ratio = bill_length_mm / bill_depth_mm) %>%
  summarize(mean_bill_ltd_ratio = mean (bill_ltd_ratio),
            median_bill_ltd_ratio = median(bill_ltd_ratio))
#Average Ratio by Group
penguins %>%
  group_by(species) %>%
  mutate(bill_ltd_ratio = bill_length_mm / bill_depth_mm) %>%
  summarize(mean_bill_ltd_ratio = mean (bill_ltd_ratio),
            median_bill_ltd_ratio = median(bill_ltd_ratio))

```

Average body mass by year

```{r}
penguins %>%
  group_by(year) %>%
  summarize(mean_body_mass = mean(body_mass_g))
```

## Data Visualization: 

-   What is the distribution of penguins flipper lengths? numerical

-   What is the distribution of penguin species? categorical

-   Does the distribution of flipper length depend of the species of penguins?

-   How does bill length change throughout the years

-   Is there any correlation between the bill length and bill depth? scatterplot

```{r}
penguins %>%
  ggplot () +
  geom_histogram( aes(x = flipper_length_mm), 
                  bins = 15, 
                  fill = "skyblue",
                  color = "red") +
  labs(title = "Distribution of Flipper length (mm)",
       subtitle = "Mean in Black, Median in Orange",
       y = "", x = "Flipper length (mm)") +
         geom_vline(aes (xintercept = mean (flipper_length_mm)), lwd = 2, lty = "dashed" ) +
          geom_vline(aes (xintercept = median (flipper_length_mm)),color = "orange", lwd = 2, lty = "dotted")
 
                     
 

```

Categorical Plot

```{r}
penguins %>%
  ggplot() +
  geom_bar(mapping = aes(x = species ), color = "black", fill="blue") +
  labs(title =  "Counts of Penguin species",
       x = "Species", y = "Count")
```

Let's make a scatter plot see if bill legnth is correlated with bill depth

```{r}
penguins %>%
  ggplot() +
  geom_point(aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_smooth(aes(x =bill_length_mm, y = bill_depth_mm, color = species), method = "lm")
```
